Density Matrix Susceptibility Tensors

### 2018-06-01 papers/1_dmatpol/paper-r1.tex
Table in paper initiated

% tables
%\usepackage{dcolumn}
%\usepackage{multirow}
%\usepackage{hhline}
%\usepackage{booktabs}
%\usepackage{longtable}

@@

Performance descriptors of the \emph{ab initio} and generalized 
models of the density matrix susceptibility are shown in Table~XXX. 
We considered three different types of point charge environment that generate
weak (0.001 -- 0.005 a.u.), 
moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
electric fields (on average) around the water molecule.
%
%for a set of varying point charge environments
%with weak, moderate and strong electric fields are shown in Table~XXX.

@@

In this Work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
at the HF level was derived for a closed shell molecule in a spatially non\hyp{}uniform
electric field distribution. It was found that the change in the density matrix element
is approximately proportional to the electric field evaluated at the distributed site associated with the 
center of charge of the 
localized molecular orbital. Subsequently, the \emph{ab initio} model served as a basis
for the formulation of the generalized model that is, in principle, not limited to the HF theory,
and takes into account higher\hyp{}order curvatures of the electric field distribution as well.
The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibility tensors
that are able to interact with the electric field, square of electric field and electric field gradients.
The particular density matrix susceptibilities are determined through the least\hyp{}squares 
linear regression with respect to the enseble of electrostatically petrurbed known density matrices.

The above theoretically developed models were then used to calculate the electric field\hyp{}induced
interaction energies, dipole moments and quadrupole moments of water molecule that
interacts with various point charge environments generating electric fields of weak, moderate
and strong magnitudes. It was found that, while the \emph{ab initio} model yelds qualitatively correct
interaction energies and dipole moments, the generalized model can reach quantitative level of accuracy
for prediction of interaction energies, dipole moments and even quadrupole moments. Although 
errors in interaction energies produced by using the generalized model 
was still at best one order of magnitude larger than the errors produced by using the
usual approach based on the distributed polarizabilities, this work demonstrated that it is possible to
accurately predict the change in the density matrix of electronic system, and the accuracy of the model
can be improved by extending the parameter space of the susceptibility tensors.

It is anticipated here that the above generalized models from Eq.~\eqref{e:final-model.General}
should be studied in the future with an emphasis 
on their further optimization, especially to reach RMSE below 1 kcal/mole
in strong and very strong electric fields. Owing the capability of the proposed approach to correctly capture the 
environment\hyp{}induced changes in the one\hyp{}particle density matrix, 
the generalized density matrix susceptibility tensors 
could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
described at correlated levels of theory, particularly including the excited state chemistry.

>>>>

% tables
%\usepackage{dcolumn}
\usepackage{multirow}
%\usepackage{hhline}
%\usepackage{booktabs}
%\usepackage{longtable}

@@

Performance descriptors of the \emph{ab initio} and generalized 
models of the density matrix susceptibility are shown in Table~\ref{t:results}. 
%
% Please add the following required packages to your document preamble:
\begin{table*}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
\textbf{}                                                                           &              & \textbf{An Initio} & \multicolumn{4}{c|}{\textbf{Generalized}}                     & \textbf{Multipole} \\ \hline
\textbf{Rank of DMS}                                                                & \textbf{Set} & \textbf{1, 0}      & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W            & 0.067 (99.80)      & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8} 
                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8} 
                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W            & 3.36 (98.70)       & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8} 
                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8} 
                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W            & 6.22 (58.72)       & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8} 
                                                                                    & M            &                    &               &               &               &               &                    \\ \cline{2-8} 
                                                                                    & S            &                    &               &               &               &               &                    \\ \hline
\end{tabular}
\end{table*}
We considered three different types of point charge environment that generate
weak (0.001 -- 0.005 a.u.), 
moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
electric fields around the water molecule (see the histograms of the electric field distribution
per atomic centre in Figure~XXX).
%
%for a set of varying point charge environments
%with weak, moderate and strong electric fields are shown in Table~XXX.

@@

In this Work, the \emph{ab initio} model of the one\hyp{}particle density matrix susceptibility tensors
at the HF level was derived for a closed shell molecule in a spatially non\hyp{}uniform
electric field distribution. It was found that the change in the density matrix element
is approximately proportional to the electric field evaluated at the distributed site associated with the 
center of charge of the 
localized molecular orbital. Subsequently, the \emph{ab initio} model served as a basis
for the formulation of the generalized model that is, in principle, not limited to the HF theory,
and takes into account higher\hyp{}order curvatures of the electric field distribution as well.
The generalized model consisted of a set of adjustable parameters defining the density matrix susceptibility tensors
that are able to interact with the electric field, square of electric field and electric field gradients.
The particular density matrix susceptibilities were determined through the least\hyp{}squares 
linear regression with respect to the training enseble of electrostatically petrurbed states.

The above theoretically developed models were then used to calculate the electric field\hyp{}induced
interaction energies, dipole moments and quadrupole moments of water molecule that
interacts with various point charge environments generating electric fields of weak, moderate
and strong magnitudes. It was found that, while the \emph{ab initio} model yelds qualitatively correct
interaction energies and dipole moments, the generalized model can reach quantitative level of accuracy
for prediction of interaction energies, dipole moments and even quadrupole moments. Although 
errors in interaction energies produced by using the generalized model 
were still at best one order of magnitude larger than the errors produced by using the
usual approach based on the distributed polarizabilities, this work demonstrated that it is possible to
accurately predict the change in the density matrix of electronic system, and the accuracy of the model
can be improved by extending the parameter space of the susceptibility tensors.

It is anticipated here that the generalized models from Eq.~\eqref{e:final-model.General}
should be studied in the future with an emphasis 
on their further optimization, especially to reach RMSE below 1 kcal/mole
in strong and very strong electric fields (0.05-0.07 a.u.). 
Owing the capability of the proposed approach to correctly capture the 
environment\hyp{}induced changes in the one\hyp{}particle density matrix, 
the generalized density matrix susceptibility tensors 
could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
described at correlated levels of theory, particularly including the excited state chemistry.
